### 2018-08-01 VV570/PROJ2/tex/report.tex
task2 done!

\title{	
\normalfont \normalsize 
\textsc{Shanghai Jiao Tong University, UM-SJTU JOINT INSTITUTE} \\ [25pt] % Your university, school and/or department name(s)
\horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
\huge Introduction to Numerical Analysis \\ HW9 \\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

@@

\section{Question 1}
	\begin{enumerate}[(a)]
		\item
			\begin{proof}
				As $\mathcal{A}$ is convex, the mean value theorem can be applied in terms of $y$ s.t.
				\begin{equation}
					\Phi(t, y_2) - \Phi(t, y_1) = (y_2 - y_1) \frac{\partial \Phi(t, y)}{\partial y}\Big|_{y=\xi} 
				\end{equation}
				where $\xi \in (y_1, y_2)$.\\
				Since the there exists $c>0$ s.t. for all $(t, y)\in \mathcal{A}$
				\begin{equation}
					\Big| \frac{\partial \Phi(t, y)}{\partial y} \Big| \leq c
				\end{equation}
				Thus
				\begin{equation}
					\begin{aligned}
						& |\Phi(t, y_2) - \Phi(t, y_1)|\\
					  = & |y_2 - y_1| \Bigg| \frac{\partial \Phi(t, y)}{\partial y}\Big|_{y=\xi} \Bigg|\\
				   \leq & c |y_2 - y_1|	
					\end{aligned}
				\end{equation}
				which implies that $\Phi(t, y)$ satisfies Lipschitz condition in $y$ on $\mathcal{A}$.
			\end{proof}

		\item 
			\begin{proof}
				Let $P_1 = (t_1, y_1)$ and $P_2 = (t_2, y_2)$. Then any point $P'$ lies on the line segment joining $P_1$ and $P_2$ can be expressed as
				\begin{equation}
					\begin{aligned}
						P & = (1-\alpha)P_1 + \alpha P_2\\
						  & = ((1-\alpha)t_1 + \alpha t_2, (1-\alpha)y_1 + \alpha y_2)\\
						  & \triangleq (t', y')
					\end{aligned}
				\end{equation} 
				where $\alpha \in [0, 1]$.\\
				It's clear that $t_1$ and $t_2$ lies between $t_0$ and $T$, and $t'$ lies between $t_1$ and $t_2$. Thus, $t'$ also lies between $t_0$ and $T$.\\
				Further, it's also clear that $-\infty < y' < +\infty$.\\
				Hence $P \in \mathcal{D}$, which implies that $\mathcal{D}$ is convex.
			\end{proof}

		\item
			\begin{proof}
				Let
				\begin{equation}
					\Phi(t, y) = \frac{4t^3y}{1+t^4}
				\end{equation}
				Then
				\begin{equation}
					\frac{\partial \Phi(t, y)}{\partial y} = \frac{4t^3}{1+t^4} =\frac{4 t}{t^2 + \frac{1}{t^2}} < \frac{4}{t^2 + \frac{1}{t^2}} < \frac{4}{2 \sqrt{t^2 \frac{1}{t^2}}} = 2
				\end{equation}
				as $t\in (0,1)$.\\ 
				which implies that $\Phi(t, y)$ satisfies a Lipschitz condition in $y$.\\
				Thus, the given IVP problem has a unique solution.
			\end{proof}

		\item 
			Definitely not recommended.\\
			As $\Phi(t, y) = 1 + y^2$, then
			\begin{equation}
				\frac{\partial \Phi}{\partial y} = 2 y \triangleq \lambda y
			\end{equation}
			Here $\lambda > 0$, and the Euler's method is not stable as the error will be amplified at each iteration step. Finally the calculation will diverge.
	\end{enumerate}

\section{Question 2}
	\begin{enumerate}[(a)]
		\item 
			As $\Phi(t, y) = arctan(y)$, then
			\begin{equation}
				\Bigg| \frac{\partial \Phi}{\partial y} \Bigg| = \frac{1}{|1+y^2|} < 1 \triangleq c
			\end{equation}
		\item 
			As $\dot{y} = \Phi(t, y) = arctan(y)$, then
			\begin{equation}
				|\ddot{y}| = \Bigg| \frac{\dot{y}}{1+y^2} \Bigg| = \frac{|arctan(y)|}{1+y^2} < \frac{\pi}{2(1+y^2)} \leq \frac{\pi}{2} 
			\end{equation}
		\item 
			As the global error is bounded by
			\begin{equation}
				|e_k| \leq \frac{\tau^*}{hc}[e^{c(t_k - t_0)} - 1]
			\end{equation}
			where $\tau^* = \max\limits_{k}|\tau_k|$, $c$ is the Lipschitz constant and is taken as $1$ as has been illustrated above.\\
			Since 
			\begin{equation}
				\begin{aligned}
					\frac{\tau^*}{h} & = \max\limits_{k}\Bigg|\frac{y(t_k) - y(t_{k-1}) - h\Phi(t_{k-1}, y(t_{k-1}))}{h}\Bigg|\\
									 & \leq \max\limits_{k}\Bigg| \frac{y(t_k) - y(t_{k-1})}{h}\Bigg| + |\Phi(t_{k-1}, y(t_{k-1}))|\\
									 & = 2 \max\limits_{k} |\Phi(t_{k-1}, y(t_{k-1}))| < \pi
				\end{aligned}
			\end{equation}
			Thus, $|e_k| < \pi[e^{(t_k - t_0)}-1]$.
	\end{enumerate}

\section{Question 3}
	\begin{enumerate}[(a)]
		\item

@@

\section{Question 6}

>>>>

\title{	
\normalfont \normalsize 
\textsc{Shanghai Jiao Tong University, UM-SJTU JOINT INSTITUTE} \\ [25pt] % Your university, school and/or department name(s)
\horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
\huge Introduction to Numerical Analysis \\ Project2 \\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

@@

\section{Task 1}
	Note the symmetry of $x_1$, for any solution $(a, b)$, $(-a, b)$ is also valid.\\
	If no special claim, the norm function for a vector is taken as the 2-norm.
	\begin{enumerate}[(a)]
		\item
			For fixed-point iteration, the iteration function $G$ is taken as
			\begin{equation}
				G \triangleq
				\begin{bmatrix}
					\sqrt{x_2}\\
					\sqrt{1-x_1^2}
				\end{bmatrix}
				=
				\begin{bmatrix}
					x_1\\
					x_2\\
				\end{bmatrix}
				=x
			\end{equation}
			The iteration converges to $(0.7862, 0.6180)$ with intial guess $x_{init} = [\frac{1}{\sqrt{2}}, \frac{1}{\sqrt{2}}]^T$ and tolerance set to $10^{-10}$. 
		\item 
			The newton iteration is implemented as mentioned in the project sheet, each time the linear system
			\begin{equation}
				J_F(x_k) w_k = y_k
			\end{equation}
			is solved with the built-in function 'linsolve' inside matlab.\\
			The iteration converges to $(0.7862, 0.6180)$ with intial guess $x_{init} = [\frac{1}{\sqrt{2}}, \frac{1}{\sqrt{2}}]^T$ and tolerance set to $10^{-10}$. 
		\item
			The input of the Broyden's method is a little bit different from that in newton or fixed-point. The initial Jacobi matrix should be provided, denoted as $A_0$. In my implementation, it is given as
			\begin{equation}
				A_0 = J_F(x_0)
			\end{equation}
			An extra calculation on $x_1$ should be done before the iteration loop as the iteration process requires both $x_0$ and $x_1$. It is approximated through newton's method in my implementation as
			\begin{equation}
				x_1 \approxeq x_0 - A_0^{-1}F(x_0)
			\end{equation}
			After all the preparation work, the iteration loop can be carried out as mentioned in the project sheet, and the key part is the application of Sherman-Morrison's formula to calculated the inverse of $A_k$ iteratively as
			\begin{equation}
				A_k^{-1} = A_{k-1}^{-1} + \frac{(s_k^T A_{k-1}^{-1} y_k)s_k^T A_{k-1}^{-1}}{s_k^T A_{k-1}^{-1} y_k}
			\end{equation}
			where $s_k = x_k - x_{k-1}$ and $y_k = F(x_k) - F(x_{k-1})$. And the iteration of $x_k$ is given as
			\begin{equation}
				x_{k+1} = x_k - A_k^{-1}F(x_k)
			\end{equation}
			The iteration converges to $(0.7862, 0.6180)$ with intial guess $x_{init} = [\frac{1}{\sqrt{2}}, \frac{1}{\sqrt{2}}]^T$ and tolerance set to $10^{-10}$. 
		\item 
			With numerical tests, newton's method out-performs the other two in terms of time cost.

\section{Task 2}
	With trival inequality analysis, the exact solution is given as $x=[0.5, 0, -\frac{\pi}{6}]$, which can be used for varification.\\
	In general, the initial value for such a non-linear system is hard to know, even the range for each component is hard to estimate. In this case, the downhill simplex method can be applied to produce a fine inital guess for further calculation like newton or quasi-newton.\\
	Numerical experiments were done to identify the difference in terms of time cost.

	\begin{center}
		\begin{tabular}{ccc}
			\hline
			\quad       & Without initialization & With initialization\\
			\hline
			fixed-point & $4.033 ms$             & $3.970 ms$\\
			neweon      & $4.661 ms$             & $0.679 ms$\\
			broyden     & $4.447 ms$             & $0.799 ms$\\
			\hline
		\end{tabular}
	\end{center}

	It can be observed from the table that the downhill simplex initialization process does help a lot to reduce time cost for further newton or broyden calculation as it provide a better initial guess. \\
	However, the time cost is almost equivalent for the fixed-point iteration, which indicates that it does not benefit from the downhill simplex initialization.

\section{Task 3}
	\begin{enumerate}[(a)]
		\item

@@
